An equivalence of compositional frameworks

### 2019-05-30 EquivCompFrameworks_20180911kc.tex
getting ready to work

%\begin{comment}
\section{Introduction}
Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed that is suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the $\textbf{apex}$ of the cospan and $a_1$ and $a_2$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means that each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout to obtain an open system whose inputs are that of the first and outputs are that of the second. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_2$};
\node (B) at (1,1) {$b^\prime$};
\node (C) at (2,0) {$a_3$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,1) {$b^\prime$};
\node (E) at (4,0) {$a_3$};
\node (F) at (2,2) {$b+_{a_2} b^\prime$};
\node (G) at (2,1) {$b+b^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_b$} (F)
(D) edge node[right]{$\psi j_{b^\prime}$} (F)
(B) edge node[above]{$j_b$} (G)
(D) edge node[above]{$j_{b^\prime}$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

@@

\subsection*{Acknowledgements}
Daniel Cicala, person that Christina emailed about lax slice category? {\chris did you use it after all?} No, I did not.
%\end{comment}

@@

@@

%\begin{comment}
\section{Appendix}

{\chris Put various stuff here, fix in the very end}
Recall that a pseudofunctor $F\colon\bicat{A}\to\bicat{B}$ between bicategories $\bicat{A}$ and $\bicat{B}$ is functorial up to coherent natural isomorphism, namely for composable arrows we have $F(g\circ f)\cong Fg\circ Ff$ and $F(1_a)\cong1_{Fa}$ satisfying standard axioms. 
Given pseudofunctors $F,G\colon \bicat{A} \to \bicat{B}$, a \define{pseudonatural transformation} $\sigma$ consists of {\chris notation below needs fixing}
\begin{itemize}
\item for each object $a \in \bicat{A}$, a morphism $\sigma_a \colon F(a) \to G(a)$ in $\bicat{B}$ 
\item for every pair of objects $a$ and $b$ of $\mathbf{A}$, we have natural isomorphisms
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\mathbf{A}(a,b)$};
\node (B) at (2,0) {$\mathbf{X}(F(a),F(b))$};
\node (C) at (0,-1) {$\mathbf{X}(G(a),G(b))$};
\node (D) at (2,-1) {$\mathbf{X}(F(a),G(b))$};
\node (E) at (1,-0.5) {$\sigma_{a,b} \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(B) edge node[right]{$(\sigma_{b})_*$} (D)
(A) edge node[left]{$G$} (C)
(C) edge node[above]{$(\sigma_a)^*$} (D);
\end{tikzpicture}
\]
where $(\sigma_a)^*$ and $(\sigma_b)_*$ are the functors induced by precomposition and postcompositon, respectively. Thus for each morphism $f \colon a \to b$ in $\mathbf{A}$, we have an invertible 2-morphism $\sigma_f \colon G(f) \sigma_a \xrightarrow{\sim} \sigma_{b} F(f)$ in $\mathbf{X}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(a)$};
\node (B) at (1,0) {$F(b)$};
\node (C) at (0,-1) {$G(a)$};
\node (D) at (1,-1) {$G(b)$};
\node (E) at (0.5,-0.5) {$\sigma_f \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F(f)$} (B)
(B) edge node[right]{$\sigma_{b}$} (D)
(A) edge node[left]{$\sigma_a$} (C)
(C) edge node[above]{$G(f)$} (D);
\end{tikzpicture}
\]
which is compatible with composition and identities.
\end{itemize}
%such that for each composable pair of morphisms $f \colon a \to b$ and $g \colon b \to c$ of $\mathbf{A}$, the following diagrams commute:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$(G(g)G(f))\sigma_a$};
%\node (A') at (2,0.5) {$G(g)(G(f) \sigma_a)$};
%\node (B) at (0,-0.5) {$G(gf) \sigma_a$};
%\node (C) at (4,0.5) {$G(g) (\sigma_b F(f))$};
%\node (C') at (4,-0.5) {$\sigma_c F(gf)$};
%\node (D) at (6,0.5) {$(G(g) \sigma_b) F(f)$};
%\node (D') at (8,-0.5) {$\sigma_c (F(g)F(f))$};
%\node (F) at (8,0.5) {$(\sigma_c F(g)) F(f)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$a^\prime$} (A')
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(A') edge node[above]{$1_{G(g)} \sigma_f$} (C)
%(B) edge node[above]{$\sigma_{gf}$} (C')
%(C) edge node [above] {${a^\prime}^{-1}$} (D)
%(D') edge node [above] {$1_{\sigma_c} \phi$} (C')
%(D) edge node [above] {$\sigma_g 1_{F(f)}$} (F)
%(F) edge node [right] {$a^\prime$} (D');
%\end{tikzpicture}
%\]
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$1_{G(a)} \sigma_a$};
%\node (B) at (1,1) {$G(1_a) \sigma_a$};
%\node (C) at (2,0) {$\sigma_a F(1_a)$};
%\node (D) at (0.5,-1) {$\sigma_a$};
%\node (E) at (1.5,-1) {$\sigma_a 1_{F(a)}$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(B) edge node[right]{$\sigma_{1_a}$} (C)
%(A) edge node [left] {$\ell^\prime$} (D)
%(D) edge node [above] {${\rho^\prime}^{-1}$} (E)
%(E) edge node [right] {$1_{\sigma_a} \phi$} (C);
%\end{tikzpicture}
%\]
%Christina: I removed the above, they are not needed I believe
We denote by $[\A,\bicat{Cat}]_\pse$ the 2-category 
of pseudofunctors, pseudonatural transformations and modifications from an ordinary category $\A$ viewed as a 2-category with trivial 2-cells, into $\bicat{Cat}$. Classicaly this has been referred to as the 2-category of \emph{opindexed categories}, since an indexed category is a contravariant pseudofunctor into $\bicat{Cat}$.

@@

		\item $S\mathbb{F}_1= \mathbb{F}_0S$ and $T\mathbb{F}_1 = \mathbb{F}_0T$ are equations between monoidal functors, and
		\item the composition and unit comparisons $\phi(N_1,N_2) \maps \mathbb{F}_1(N_1) \odot \mathbb{F}_1(N_2) \to \mathbb{F}_1(N_1\odot N_2)$ and $\phi_A \maps U_{\mathbb{F}_0 (A)} \to \mathbb{F}_1(U_A)$ are monoidal natural transformations.
	\end{itemize}
The monoidal lax double functor is \define{braided} if $\mathbb{F}_0$ and $\mathbb{F}_1$ are braided monoidal functors and \define{symmetric} if they are symmetric monoidal functors, and lax monoidal or oplax monoidal if the isomorphisms and families of natural isomorphisms of items (1)-(4) above are merely morphisms and natural transformations going in the appropriate directions.
\end{defn}
%\end{comment}

@@

>>>>

\section{Introduction}
\begin{comment}
Networks are playing an increasingly prominent role in our understanding of the world and as a result, methods of characterizing and studying networks have become evermore necessary. Applied category theory provides such an avenue and much work has been done towards this effort  \cite{BC,BCR,BF,BFP,BP,Cour,Fong,JM,LS,Yass}. One of the more recent frameworks developed that is suitable for studying networks is Fong's `decorated cospans' \cite{Fong}. A $\textbf{cospan}$ in any category is a diagram of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
We call $b$ the $\textbf{apex}$ of the cospan and $a_1$ and $a_2$ the $\textbf{feet}$ of the cospan. The morphisms $i$ and $o$ are called the $\textbf{legs}$. Cospans are ideal for realizing networks and more generally `open' systems where here open means that each system or network comes with prescribed inputs and outputs which are represented by the feet of the cospan $a$ and $b$, respectively. Two open systems viewed as cospans such that the outputs of the first and the inputs of the second coincide can then naturally be composed via pushout to obtain an open system whose inputs are that of the first and outputs are that of the second. For example, we can compose the above cospan with the following:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_2$};
\node (B) at (1,1) {$b^\prime$};
\node (C) at (2,0) {$a_3$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i^\prime$} (B)
(C) edge node[above]{$o^\prime$} (B);
\end{tikzpicture}
\]
to obtain a new cospan whose left foot is the left foot of the first, whose right foot is the right foot of the second, and whose apex and legs are given by the pushout.
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a_1$};
\node (B) at (1,1) {$b$};
\node (C) at (2,0) {$a_2$};
\node (D) at (3,1) {$b^\prime$};
\node (E) at (4,0) {$a_3$};
\node (F) at (2,2) {$b+_{a_2} b^\prime$};
\node (G) at (2,1) {$b+b^\prime$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(C) edge node[above]{$i^\prime$} (D)
(E) edge node[above]{$o^\prime$} (D)
(B) edge node[left]{$\psi j_b$} (F)
(D) edge node[right]{$\psi j_{b^\prime}$} (F)
(B) edge node[above]{$j_b$} (G)
(D) edge node[above]{$j_{b^\prime}$} (G)
(G) edge node [left] {$\psi$} (F);
\end{tikzpicture}
\]

@@

\subsection*{Acknowledgements}
Daniel Cicala, various others.
\end{comment}

@@

Details of the constructions and the proof of the above theorem can be found at \cite{CourPhD}.

@@

\section{Appendix}

\begin{comment}
{\chris Put various stuff here, fix in the very end}
Recall that a pseudofunctor $F\colon\bicat{A}\to\bicat{B}$ between bicategories $\bicat{A}$ and $\bicat{B}$ is functorial up to coherent natural isomorphism, namely for composable arrows we have $F(g\circ f)\cong Fg\circ Ff$ and $F(1_a)\cong1_{Fa}$ satisfying standard axioms. 
Given pseudofunctors $F,G\colon \bicat{A} \to \bicat{B}$, a \define{pseudonatural transformation} $\sigma$ consists of {\chris notation below needs fixing}
\begin{itemize}
\item for each object $a \in \bicat{A}$, a morphism $\sigma_a \colon F(a) \to G(a)$ in $\bicat{B}$ 
\item for every pair of objects $a$ and $b$ of $\mathbf{A}$, we have natural isomorphisms
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$\mathbf{A}(a,b)$};
\node (B) at (2,0) {$\mathbf{X}(F(a),F(b))$};
\node (C) at (0,-1) {$\mathbf{X}(G(a),G(b))$};
\node (D) at (2,-1) {$\mathbf{X}(F(a),G(b))$};
\node (E) at (1,-0.5) {$\sigma_{a,b} \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F$} (B)
(B) edge node[right]{$(\sigma_{b})_*$} (D)
(A) edge node[left]{$G$} (C)
(C) edge node[above]{$(\sigma_a)^*$} (D);
\end{tikzpicture}
\]
where $(\sigma_a)^*$ and $(\sigma_b)_*$ are the functors induced by precomposition and postcompositon, respectively. Thus for each morphism $f \colon a \to b$ in $\mathbf{A}$, we have an invertible 2-morphism $\sigma_f \colon G(f) \sigma_a \xrightarrow{\sim} \sigma_{b} F(f)$ in $\mathbf{X}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$F(a)$};
\node (B) at (1,0) {$F(b)$};
\node (C) at (0,-1) {$G(a)$};
\node (D) at (1,-1) {$G(b)$};
\node (E) at (0.5,-0.5) {$\sigma_f \Nearrow$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$F(f)$} (B)
(B) edge node[right]{$\sigma_{b}$} (D)
(A) edge node[left]{$\sigma_a$} (C)
(C) edge node[above]{$G(f)$} (D);
\end{tikzpicture}
\]
which is compatible with composition and identities.
\end{itemize}
%such that for each composable pair of morphisms $f \colon a \to b$ and $g \colon b \to c$ of $\mathbf{A}$, the following diagrams commute:
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0.5) {$(G(g)G(f))\sigma_a$};
%\node (A') at (2,0.5) {$G(g)(G(f) \sigma_a)$};
%\node (B) at (0,-0.5) {$G(gf) \sigma_a$};
%\node (C) at (4,0.5) {$G(g) (\sigma_b F(f))$};
%\node (C') at (4,-0.5) {$\sigma_c F(gf)$};
%\node (D) at (6,0.5) {$(G(g) \sigma_b) F(f)$};
%\node (D') at (8,-0.5) {$\sigma_c (F(g)F(f))$};
%\node (F) at (8,0.5) {$(\sigma_c F(g)) F(f)$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[above]{$a^\prime$} (A')
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(A') edge node[above]{$1_{G(g)} \sigma_f$} (C)
%(B) edge node[above]{$\sigma_{gf}$} (C')
%(C) edge node [above] {${a^\prime}^{-1}$} (D)
%(D') edge node [above] {$1_{\sigma_c} \phi$} (C')
%(D) edge node [above] {$\sigma_g 1_{F(f)}$} (F)
%(F) edge node [right] {$a^\prime$} (D');
%\end{tikzpicture}
%\]
%\[
%\begin{tikzpicture}[scale=1.5]
%\node (A) at (0,0) {$1_{G(a)} \sigma_a$};
%\node (B) at (1,1) {$G(1_a) \sigma_a$};
%\node (C) at (2,0) {$\sigma_a F(1_a)$};
%\node (D) at (0.5,-1) {$\sigma_a$};
%\node (E) at (1.5,-1) {$\sigma_a 1_{F(a)}$};
%\path[->,font=\scriptsize,>=angle 90]
%(A) edge node[left]{$\psi 1_{\sigma_a}$} (B)
%(B) edge node[right]{$\sigma_{1_a}$} (C)
%(A) edge node [left] {$\ell^\prime$} (D)
%(D) edge node [above] {${\rho^\prime}^{-1}$} (E)
%(E) edge node [right] {$1_{\sigma_a} \phi$} (C);
%\end{tikzpicture}
%\]
%Christina: I removed the above, they are not needed I believe
We denote by $[\A,\bicat{Cat}]_\pse$ the 2-category 
of pseudofunctors, pseudonatural transformations and modifications from an ordinary category $\A$ viewed as a 2-category with trivial 2-cells, into $\bicat{Cat}$. Classicaly this has been referred to as the 2-category of \emph{opindexed categories}, since an indexed category is a contravariant pseudofunctor into $\bicat{Cat}$.

@@

		\item $S\mathbb{F}_1= \mathbb{F}_0S$ and $T\mathbb{F}_1 = \mathbb{F}_0T$ are equations between monoidal functors, and
		\item the composition and unit comparisons $\phi(N_1,N_2) \maps \mathbb{F}_1(N_1) \odot \mathbb{F}_1(N_2) \to \mathbb{F}_1(N_1\odot N_2)$ and $\phi_A \maps U_{\mathbb{F}_0 (A)} \to \mathbb{F}_1(U_A)$ are monoidal natural transformations.
	\end{itemize}
The monoidal lax double functor is \define{braided} if $\mathbb{F}_0$ and $\mathbb{F}_1$ are braided monoidal functors and \define{symmetric} if they are symmetric monoidal functors, and lax monoidal or oplax monoidal if the isomorphisms and families of natural isomorphisms of items (1)-(4) above are merely morphisms and natural transformations going in the appropriate directions.
\end{defn}
\end{comment}

@@

\bibitem{CourPhD} K.\ Courser, Cospans for Networks, PhD Thesis, University of California at Riverside (2019)
